Non-kissing and non-crossing complexes \\ for locally gentle algebras

### 2018-07-11 draft/accordionComplexSurfaces.tex
Minor change
changes phone

\begin{definition}%[\cite{McConville, PaluPilaudPlamondon}]
\label{def:nKc}
The \defn{non-kissing complex} of~$\bar Q$ is the simplicial complex~$\NKC$ whose faces are the collections of pairwise non-kissing walks of~$\NKWalks^\pm(\bar Q)$.
Note that self-kissing walks never appear in~$\NKC$ by definition.
In contrast, no straight walk can kiss another walk by definition, so that they appear in all facets of~$\NKC$.
The \defn{reduced non-kissing complex}~$\RNKC$ is the simplicial complex whose faces are the collections of pairwise non-kissing bended walks of~$\bar Q$.
\end{definition}

@@

The following definition generalises the one from \cite{MannevillePilaud-accordion} for the case where~$\bar \surface$ is a disk.  It appears in \cite{BaurSimoes} under the name ``permissible arc''.
\pierreguy{Is it ok to cite like this?}

@@

In this section, we show that non-kissing complexes and non-crossing complexes actually coincide.
Namely, any pair of dual cellular dissections~$(\dissection, \dissection\dual)$ defines a locally gentle bound quiver~$\bar Q_{\dissection, \dissection\dual}$ such that the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
Conversely, any locally gentle bound quiver~$\bar Q$ gives rise to a surface~$\surface_{\bar Q}$ equipped with a pair of dual cellular dissections~$\dissection_{\bar Q}$ and~${\dissection_{\bar Q}}\dual$ such that the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$.
In fact, we show that the accordion, slalom, and non-kissing complexes can all be pictured in a unique way on the surface.
Interestingly, via the correspondence between pairs of dual cellular dissections and locally gentle bound quivers, the duality between the dissections~$\dissection$ and~$\dissection\dual$ translates to the Koszul duality between the locally gentle algebras of~$\bar Q$ and~$\bar Q\dual$.

@@

\begin{proof}
 The duality gives a bijection~$v\in (Q_\dissection)_0\mapsto v\dual\in (Q_{\dissection\dual})_0$, and induces a bijection~$\alpha\in (Q_\dissection)_1\mapsto \alpha\dual\in (Q_{\dissection\dual})_1$ with~$s(\alpha\dual)=s(\alpha)\dual$ and~$t(\alpha\dual)=t(\alpha)\dual$.
 Moreover, for any two composable arrows~$\alpha,\beta\in(Q_\dissection)_0$, we have~$\alpha\beta\in I_\dissection$ if and only if~$s(\alpha),t(\alpha),t(\beta)$ are three consecutive edges of a~$\dissection$-cell~$f$, if and only if~$s(\alpha)\dual,t(\alpha)\dual,t(\beta)\dual$ are three consecutive arcs in~$\dissection\dual$ ending in~$f\dual$, if and only if~$\beta\dual\alpha\dual\notin I_{\dissection\dual}$.
\end{proof}

@@

We now associate a surface to a locally gentle bound quiver.
This construction is directly inspired by the one found in \cite{OpperPlamondonSchroll} (see Remark \ref{rem:comparisonWithOPS} for a comparison of the two constructions).

\begin{definition}
\label{def:surfaceQuiver}
The surface~$\surface(\bar Q)$ of a locally gentle bound quiver~$\bar Q = (Q,I)$ is the surface obtained from the blossoming quiver~$\bar Q\blossom$ as follows:
\begin{enumerate}[(i)]
\item for each arrow~$\alpha \in Q_1\blossom$, consider a lozange~$L(\alpha)$ with four (oriented) sides
\begin{alignat*}{2}
& \Enrs{\alpha} = [v(\alpha), s(\alpha)] \qquad & \Enrt{\alpha} = [v(\alpha), t(\alpha)] \qquad & \text{(colored green)} \\
\text{and} \qquad & \Ers{\alpha} = [v\dual(\alpha), s(\alpha)] \qquad & \Ert{\alpha} = [v\dual(\alpha), t(\alpha)] \qquad & \text{(colored red)}
\end{alignat*}
placed as follows:

@@

Definition~\ref{def:surfaceQuiver} constructs an orientable surface~$\surface(\bar Q)$ with boundaries.
\fref{fig:surfaces} illustrates this construction for the quivers of \fref{fig:quivers}.

@@

The advantage of Definition~\ref{def:surfaceQuiver} is that it automatically endows with two disjoint sets~$V_{\bar Q}$ and~${V_{\bar Q}}\dual$ of marked points and two dual cellular dissections~$\dissection_{\bar Q}$ and~${\dissection_{\bar Q}}\dual$ defined as follows.

\begin{definition}
\label{def:dissectionQuiver}
The surface~$\surface(\bar Q)$ is endowed with
\begin{itemize}
\item the set~$V_{\bar Q}$ of points~$v(\alpha)$ for~$\alpha \in Q_1\blossom$ after the identifications given by~(ii),
\item the $V_{\bar Q}$-dissection~$\dissection_{\bar Q}$ given by all sides~$\Enrs{\alpha}$ and~$\Enrt{\alpha}$ for~$\alpha \in Q_1\blossom$ after the identifications given by~(ii).
\end{itemize}
The set~${V_{\bar Q}}\dual$ and the ${V_{\bar Q}}\dual$-dissection~${\dissection_{\bar Q}}\dual$ are defined dually.
\end{definition}

\begin{proposition}
\label{prop:dissectionsAreCellular}
Let~$\bar Q$ be a locally gentle bound quiver.
Then the dissections~$\dissection_{\bar Q}$ and~$\dissection_{\bar Q}\dual$ are cellular and dual to each other.
\end{proposition}
\begin{proof}
In the construction of~$\surface(\bar Q)$, if two arrows~$\alpha$ and~$\beta$ of~$\bar Q \blossom$ are such that~$t(\alpha) = s(\beta)$ and~$\alpha\beta\in I\blossom$, then the vertices~$v(\alpha)$ and~$v(\beta)$ are identified on the surface.
Thus, any two arrows of a given path contribute the same vertex of~$V_{\bar Q}$, so~$V_{\bar Q}$ is in bijection with the set of maximal paths in~$\bar Q\blossom$.

@@

In both cases, we see that~$v$ is enclosed in a polygon of~$\dissection_{\bar Q}\dual$.  This shows that~$\dissection_{\bar Q}\dual$ is a cellular dissection of~$\surface(\bar Q)$.
Dually, one shows that~$\dissection_{\bar Q}$ is also a cellular dissection.

Moreover, by construction, each edge of~$\dissection_{\bar Q}$ which is not on the boundary of~$\surface(\bar Q)$ crosses exactly one edge of~$\dissection_{\bar Q}\dual$, and vice versa.
This shows that~$\dissection_{\bar Q}$ and~$\dissection_{\bar Q}\dual$ are dual to each other.
\end{proof}

@@

\begin{remark}
\label{rem:propertiesSurface}
The following observations are useful for the computation of examples.
\begin{enumerate}[(i)]
\item The set~$V_{\bar Q}$ has one vertex for each straight walk in~$\bar Q$ (equivalently, for each maximal path in~$\bar Q$).
      Finite straight walks yield vertices on the boundary, while infinite cyclic straight walks in~$\bar Q$ and in~$\bar Q\koszul$ yield punctures.
\item \label{item:arcs}
      The dissection~$\dissection_{\bar Q}$ has one arc for each vertex~$p \in \bar Q_0$, obtained by concataining the sides~$\Enrt{\alpha} = \Enrs{\beta}$ and~$\Enrt{\alpha'} = \Enrs{\beta'}$ where~$p = t(\alpha) = s(\beta) = t(\alpha') = s(\beta')$, $\alpha\beta \notin I$ and~$\alpha'\beta' \notin I$.
\item The dissection~$\dissection_{\bar Q}$ has one $\ell$-cell for each straight walk of length~$\ell$ in~$\bar Q\koszul$.
\item A similar remark holds dually for~${V_{\bar Q}}\dual$ and~${\dissection_{\bar Q}}\dual$.
\item The number~$p$ of punctures of~$\surface(\bar Q)$ is the number of infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$.
\item \label{item:boundary}
      The number~$b$ of boundary components of~$\surface(\bar Q)$ can be computed as follows.
      There are two natural perfect matchings whose vertices are the blossom vertices of~$\bar Q$: one is obtained by joining the endpoints of each finite straight walk of~$\bar Q$, and the other is obtained similarly from~$\bar Q \koszul$.
      Let~$G$ be the superposition of these two perfect matchings.
      Then the number~$b$ of boundary components of~$\surface(\bar Q)$ is the number of connected components of~$G$.
\item The genus of the surface~$\surface(\bar Q)$ is
      \[
       g = \frac{|Q_1| - |Q_0| - b - p + 2}{2},
      \]
      where~$b$ is the number of boundary components (see above for a way to compute~$b$) and~$p$ the number of punctures (\ie infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$).
      Indeed, $2g = 2-\chi(\surface(\bar Q))$, where~$\chi(\surface(\bar Q))$ is the Euler characteristic of~$\surface(\bar Q)$.
      The dissection~$\dissection$ provides a cellular decomposition of~$\surface(\bar Q)$ on which we fill the boundary components and the punctures with disks.  
      The number of faces of this cellular decomposition is~$b$, its number of edges is~$|Q_0|$ and its number of vertices is the number of finite straight walks in~$\bar Q$, which is equal to~$2|Q_0| - |Q_1|$.
      The above formula for~$g$ follows.
\end{enumerate}
\end{remark}

@@

@@

Let~$\bar Q$ be a locally gentle bound quiver.
For each arc of the dissection~$\dissection_{\bar Q}$ on~$\surface(\bar Q)$, we fix a point on the interior of this arc, which we call its ``middle point''.
To each walk on~$\bar Q$, we will now associate a~$\dissection_{\bar Q}$-accordion.

\begin{definition}
\label{def:curveOfAnArrow}
For any arrow~$\alpha$ on $\bar Q\blossom$, let~$\curveof(\alpha)$ be the curve on~$\surface(\bar Q)$ which goes from the middle point of the arc corresponding to $s(\alpha)$ to the middle point of the arc corresponding to $t(\alpha)$ by following the angle corresponding to~$\alpha$.
Define~$\curveof(\alpha^{-1})$ to be~$\curveof(\alpha)^{-1}$.
\end{definition}

@@

\begin{proposition}
\label{prop:walks=arcs}
The maps~$\curveof(-)$ and~$\walk(-)$ induce mutually inverse bijections between the set of undirected walks on~$\bar Q$ and the set of~$\dissection_{\bar Q}$-accordions on~$\surface(\bar Q)$.
\end{proposition}

\begin{lemma}
\label{lem:nonKissing=nonCrossing}
Let~$\omega_1$ and~$\omega_2$ be walks on~$\bar Q$.  Then~$\omega_1$ and~$\omega_2$ are non-kissing if and only if the~$\dissection_{\bar Q}$-accordions~$\curveof(\omega_1)$ and~$\curveof(\omega_2)$ are non-crossing on~$\surface(\bar Q)$.
\end{lemma}

@@

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=1]{kissings}}
	\caption{Two representations of a pair of kissing walks.}
	\label{fig:kissings}
\end{figure}
The left hand side of the picture is as in \fref{fig:kissing}, and the right hand side is simply a different representation of it.
On the surface~$\surface(\bar Q)$, the dual dissections~$\dissection$ and~$\dissection\dual$ are as on \fref{fig:kissingVSCrossing} (on the right hand side).

@@

\begin{example}
\label{exm:peakDeepFacets}
\vincent{TODO: Describe here the peak and deep facets in terms of the surface.}
\end{example}

@@

 \para{Geometric proof}
 Let~$F$ be a face of the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$.
 Its elements are pairwise compatible~$\dissection$-accordions on the surface~$\surface(\bar Q)$.
 Thus~$F$ can be viewed as a partial tagged triangulation of~$\surface(\bar Q)$, by letting the~$\dissection$-accordions circling a puncture in (say) clockwise order be seen as ``notched'' arcs to said puncture, and the ones circling it in counter-clockwise orientation be ``un-notched'' arcs.
 It is known (see, for instance, \cite{FominShapiroThurston}) that any such partial tagged triangulation can be embedded into a full triangulation, and that triangulations contain a finite number of arcs.
 Thus~$F$ is finite, and is contained in a finite maximal face.
\end{proof}

@@

\para{Combinatorial proof}
Consider a maximal non-kissing face~$F \in \NKC$ and an arrow~$\alpha$ in~$c$.
Let~$\beta, \gamma \in Q_1\blossom$ be such that~$t(\alpha) = s(\beta) = s(\gamma)$ and~$\alpha\beta \notin I$ while~$\alpha\gamma \in I$ (in other words, $\beta$ is the next arrow along~$c$ after~$\alpha$ and~$\gamma$ is the other outgoing arrow at~$t(\alpha)$).
Consider the walk~$\mu \eqdef \distinguishedWalk{\gamma}{F}$ and decompose it as~$\mu = \mu_1 \gamma \mu_2$ according to this occurence of~$\gamma$ on~$\mu$.
Consider the walk~$\nu \eqdef \mu_2^{-1} \gamma^{-1} c^\infty$.
If~$\nu$ is contained in~$F$, then we have found a walk of the desired form (remember that walks are considered as undirected).
Otherwise, there exists a walk~$\lambda$ of~$F$ preventing~$\nu$ to belong to~$F$.
Let~$\sigma$ be a common substring of~$\nu$ and~$\lambda$ where they kiss.
Since~$\lambda$ must be compatible with~$\mu$ and with the straight walk~$c^\infty$, the substring~$\sigma$ must contain~$t(\alpha) = s(\beta)$.
If~$\lambda$ contains neither~$\alpha$ nor~$\beta$, then it contains~$\gamma$ and enters $\nu$ at~$s(\gamma)$.
Since $\lambda$ kisses~$\nu$, it also enters~$\mu_2$, contradicting the maximality of~$\mu$ for~$\prec_\gamma$.
Therefore, $\lambda$ contains either~$\alpha$ or~$\beta$.
We have thus found a bended walk~$\lambda \in F$ containing an arrow~$\alpha$ or~$\beta$ of~$c$, say~$\alpha$.
We can now moreover assume that~$\lambda = \distinguishedWalk{\alpha}{F}$.
If~$\lambda$ starts by~$^\infty c$, or ends by~$c^\infty$, then we have found a walk of the desired form.
Otherwise, decompose this walk as~$\lambda = \lambda_1 \alpha \lambda_2$ and consider the walk~$\omega = \lambda_1 c^\infty$.
We claim that~$\omega$ belongs to~$F$.
Otherwise, there exists a walk~$\rho$ preventing~$\omega$ to belong to~$F$.
Let~$\tau$ be a common substring of~$\omega$ and~$\rho$ where they kiss.
Since~$\omega$ must be compatible with~$\lambda$ and with the straight walk~$c^\infty$, the substring~$\tau$ must strictly contain the substring of~$\lambda$ along the cycle~$c$.
It follows that $\rho$ leaves~$\omega$ when it is on the cycle~$c$, thus with an incoming arrow, contradicting the maximality of~$\lambda$ for~$\prec_\alpha$.

\para{Geometric proof}
The statement of the lemma translates on~$\surface(\bar Q)$ as follows: for any facet~$F$ of the non-crossing complex~$\NCC$ and any puncture of the dissection~$\dissection_{\bar Q}$, there exists a curve in~$F$ which goes to the puncture.
Assume that there is a puncture with no curves of~$F$ going to it.
Construct a curve~$\gamma$ as follows:~$\gamma$ first unwinds from the puncture, then~$\gamma$ leaves the local picture around the puncture by turning around another marked point (marked~$X$ on the picture below).

@@

%\begin{proposition}
%\label{prop:gcdVectorsSurface}
%For any walk~$\omega$ in any non-kissing facet~$F \in \NKC$,
%\begin{itemize}
%\item the~$\b{d}$-vectors of the walk~$\omega$, of the $\dissection$-accordion~$\curveof(\omega)$, and of the $\dissection\dual$-slalom~$\curveof(\omega)$ coincide,
%\item the~$\b{g}$-vectors of the walk~$\omega$, of the $\dissection$-accordion~$\curveof(\omega)$, and of the $\dissection\dual$-slalom~$\curveof(\omega)$ coincide,
%\item the~$\b{c}$-vectors of the walk~$\omega$ in~$F$, of the $\dissection$-accordion~$\curveof(\omega)$ in~$\curveof(F)$ and of the $\dissection\dual$-slalom~$\curveof(\omega)$ in~$\curveof(F)$ concide.
%\end{itemize}
%\end{proposition}
%
%\begin{proof}
%Definitions~\ref{def:cVectorsAccordions} and~\ref{def:gVectorsSlaloms} are immediate translations in terms of curves on the surface~$\surface(\bar Q)$ of Definitions~\ref{def:cVectorsWalks} and~\ref{def:gVectorsWalks} respectively.
%\end{proof}

>>>>

\begin{definition}%[\cite{McConville, PaluPilaudPlamondon}]
\label{def:nKc}
The \defn{non-kissing complex} of~$\bar Q$ is the simplicial complex~$\NKC$ whose faces are the collections of pairwise non-kissing walks of~$\NKWalks^\pm(\bar Q)$.
Note that self-kissing walks never appear in~$\NKC$ by definition.
In contrast, no straight walk can kiss another walk by definition, so that they appear in all facets of~$\NKC$.
The \defn{reduced non-kissing complex}~$\RNKC$ is the simplicial complex whose faces are the collections of pairwise non-kissing bending walks of~$\bar Q$.
\end{definition}

@@

The following definition generalises the one from \cite{MannevillePilaud-accordion} for the case where~$\bar \surface$ is a disk.  It appears in \cite{BaurSimoes} under the name ``permissible arc''.

@@

In this section, we show that non-kissing and non-crossing complexes actually coincide.
Namely, any pair of dual cellular dissections~$(\dissection, \dissection\dual)$ defines a locally gentle bound quiver~$\bar Q_{\dissection, \dissection\dual}$ such that the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
Conversely, any locally gentle bound quiver~$\bar Q$ gives rise to a surface~$\surface_{\bar Q}$ equipped with a pair of dual cellular dissections~$\dissection_{\bar Q}$ and~${\dissection_{\bar Q}}\dual$ such that the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$.
In fact, we show that the accordion, slalom, and non-kissing complexes can all be pictured in a unique way on the surface.
Interestingly, via the correspondence between pairs of dual cellular dissections and locally gentle bound quivers, the duality between the dissections~$\dissection$ and~$\dissection\dual$ translates to the Koszul duality between the locally gentle algebras of~$\bar Q$ and~$\bar Q\koszul$.

@@

\begin{proof}
The duality provides a bijection~$v\in (Q_\dissection)_0\mapsto v\dual\in (Q_{\dissection\dual})_0$, and thus induces a bijection~${\alpha \in (Q_\dissection)_1 \mapsto \alpha\dual \in (Q_{\dissection\dual})_1}$ with~$s(\alpha\dual)=s(\alpha)\dual$ and~$t(\alpha\dual)=t(\alpha)\dual$.
 Moreover, for any two composable arrows~$\alpha,\beta\in(Q_\dissection)_0$, we have~$\alpha\beta\in I_\dissection$ if and only if~$s(\alpha),t(\alpha),t(\beta)$ are three consecutive edges of a~$\dissection$-cell~$f$, if and only if~$s(\alpha)\dual,t(\alpha)\dual,t(\beta)\dual$ are three consecutive arcs in~$\dissection\dual$ ending in~$f\dual$, if and only if~$\beta\dual\alpha\dual\notin I_{\dissection\dual}$.
\end{proof}

@@

We now associate a surface to a locally gentle bound quiver.
This construction yields the same surface as the one constructed in \cite{OpperPlamondonSchroll} (see Remark \ref{rem:comparisonWithOPS} for a comparison of the two constructions).

\begin{definition}
\label{def:surfaceQuiver}
The surface~$\surface_{\bar Q}$ of a locally gentle bound quiver~$\bar Q = (Q,I)$ is the surface obtained from the blossoming quiver~$\bar Q\blossom$ as follows:
\begin{enumerate}[(i)]
\item for each arrow~$\alpha \in Q_1\blossom$, consider a lozange~$L(\alpha)$ with four (oriented) sides
\begin{alignat*}{2}
& \Enrs{\alpha} = [v(\alpha), s(\alpha)] \qquad & \Enrt{\alpha} = [v(\alpha), t(\alpha)] \qquad & \text{(colored green)} \\
\text{and} \qquad & \Ers{\alpha} = [v\dual(\alpha), s(\alpha)] \qquad & \Ert{\alpha} = [v\dual(\alpha), t(\alpha)] \qquad & \text{(colored red)}
\end{alignat*}
placed as follows:

@@

Definition~\ref{def:surfaceQuiver} constructs an orientable surface~$\surface_{\bar Q}$ with boundaries.
\fref{fig:surfaces} illustrates this construction for the quivers of \fref{fig:quivers}.

@@

The advantage of Definition~\ref{def:surfaceQuiver} is that it automatically endows~$\surface_{\bar Q}$ with two disjoint sets~$V_{\bar Q}$ and~${V_{\bar Q}}\dual$ of marked points and two dual cellular dissections~$\dissection_{\bar Q}$ and~${\dissection_{\bar Q}}\dual$ defined as follows.

\begin{definition}
\label{def:dissectionQuiver}
The surface~$\surface_{\bar Q}$ is endowed with
\begin{itemize}
\item the set~$V_{\bar Q}$ of points~$v(\alpha)$ for~$\alpha \in Q_1\blossom$ after the identifications given by~(ii),
\item the $V_{\bar Q}$-dissection~$\dissection_{\bar Q}$ given by all sides~$\Enrs{\alpha}$ and~$\Enrt{\alpha}$ for~$\alpha \in Q_1\blossom$ after the identifications given by~(ii).
\end{itemize}
The set~${V_{\bar Q}}\dual$ and the ${V_{\bar Q}}\dual$-dissection~${\dissection_{\bar Q}}\dual$ are defined dually.
\end{definition}

\begin{proposition}
\label{prop:dissectionsAreCellular}
Let~$\bar Q$ be a locally gentle bound quiver.
Then the dissections~$\dissection_{\bar Q}$ and~$\dissection_{\bar Q}\dual$ are cellular and dual to each other.
\end{proposition}
\begin{proof}
In the construction of~$\surface_{\bar Q}$, if two arrows~$\alpha$ and~$\beta$ of~$\bar Q \blossom$ are such that~${t(\alpha) = s(\beta)}$ and~${\alpha\beta \in I\blossom}$, then the vertices~$v(\alpha)$ and~$v(\beta)$ are identified on the surface.
Thus, any two arrows of a given path contribute the same vertex of~$V_{\bar Q}$, so~$V_{\bar Q}$ is in bijection with the set of maximal paths in~$\bar Q\blossom$.

@@

In both cases, we see that~$v$ is enclosed in a polygon of~$\dissection_{\bar Q}\dual$.  This shows that~$\dissection_{\bar Q}\dual$ is a cellular dissection of~$\surface_{\bar Q}$.
Dually, one shows that~$\dissection_{\bar Q}$ is also a cellular dissection.

Moreover, by construction, each edge of~$\dissection_{\bar Q}$ which is not on the boundary of~$\surface_{\bar Q}$ crosses exactly one edge of~$\dissection_{\bar Q}\dual$, and vice versa.
This shows that~$\dissection_{\bar Q}$ and~$\dissection_{\bar Q}\dual$ are dual to each other.
\end{proof}

@@

\begin{remark}
\label{rem:propertiesSurface}
The following observations are useful for the computation of examples.
\begin{enumerate}[(i)]
\item The set~$V_{\bar Q}$ has one vertex for each straight walk in~$\bar Q$ (equivalently, for each maximal path in~$\bar Q$).
      Finite straight walks yield vertices on the boundary, while infinite cyclic straight walks in~$\bar Q$ and in~$\bar Q\koszul$ yield punctures.
\item \label{item:arcs}
      The dissection~$\dissection_{\bar Q}$ has one arc for each vertex~$p \in \bar Q_0$, obtained by concataining the sides~$\Enrt{\alpha} = \Enrs{\beta}$ and~$\Enrt{\alpha'} = \Enrs{\beta'}$ where~$p = t(\alpha) = s(\beta) = t(\alpha') = s(\beta')$, $\alpha\beta \notin I$ and~$\alpha'\beta' \notin I$.
\item The dissection~$\dissection_{\bar Q}$ has one $\ell$-cell for each straight walk of length~$\ell$ in~$\bar Q\koszul$.
\item A similar remark holds dually for~${V_{\bar Q}}\dual$ and~${\dissection_{\bar Q}}\dual$.
\item The number~$p$ of punctures of~$\surface_{\bar Q}$ is the number of infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$.
\item \label{item:boundary}
      The number~$b$ of boundary components of~$\surface_{\bar Q}$ can be computed as follows.
      There are two natural perfect matchings whose vertices are the blossom vertices of~$\bar Q$: one is obtained by joining the endpoints of each finite straight walk of~$\bar Q$, and the other is obtained similarly from~$\bar Q \koszul$.
      Let~$G$ be the superposition of these two perfect matchings.
      Then the number~$b$ of boundary components of~$\surface_{\bar Q}$ is the number of connected components of~$G$.
\item The genus of the surface~$\surface_{\bar Q}$ is
      \[
       g = \frac{|Q_1| - |Q_0| - b - p + 2}{2},
      \]
      where~$b$ is the number of boundary components (see above for a way to compute~$b$) and~$p$ the number of punctures (\ie infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$).
      Indeed, $2g = 2-\chi(\surface_{\bar Q})$, where~$\chi(\surface_{\bar Q})$ is the Euler characteristic of~$\surface_{\bar Q}$.
      The dissection~$\dissection$ provides a cellular decomposition of~$\surface_{\bar Q}$ on which we fill the boundary components and the punctures with disks.  
      The number of faces of this cellular decomposition is~$b$, its number of edges is~$|Q_0|$ and its number of vertices is the number of finite straight walks in~$\bar Q$, which is equal to~$2|Q_0| - |Q_1|$.
      The above formula for~$g$ follows.
\end{enumerate}
\end{remark}

@@

We conclude this section with four families of examples.

@@

Let~$\bar Q$ be a locally gentle bound quiver.
For each arc of the dissection~$\dissection_{\bar Q}$ on~$\surface_{\bar Q}$, we fix a point on the interior of this arc, which we call its ``middle point''.
To each walk on~$\bar Q$, we will now associate a~$\dissection_{\bar Q}$-accordion.

\begin{definition}
\label{def:curveOfAnArrow}
For any arrow~$\alpha$ on $\bar Q\blossom$, let~$\curveof(\alpha)$ be the curve on~$\surface_{\bar Q}$ which goes from the middle point of the arc corresponding to $s(\alpha)$ to the middle point of the arc corresponding to $t(\alpha)$ by following the angle corresponding to~$\alpha$.
Define~$\curveof(\alpha^{-1})$ to be~$\curveof(\alpha)^{-1}$.
\end{definition}

@@

\begin{proposition}
\label{prop:walks=arcs}
The maps~$\curveof(-)$ and~$\walk(-)$ induce mutually inverse bijections between the set of undirected walks on~$\bar Q$ and the set of~$\dissection_{\bar Q}$-accordions on~$\surface_{\bar Q}$.
\end{proposition}

\begin{lemma}
\label{lem:nonKissing=nonCrossing}
Let~$\omega_1$ and~$\omega_2$ be walks on~$\bar Q$.  Then~$\omega_1$ and~$\omega_2$ are non-kissing if and only if the~$\dissection_{\bar Q}$-accordions~$\curveof(\omega_1)$ and~$\curveof(\omega_2)$ are non-crossing on~$\surface_{\bar Q}$.
\end{lemma}

@@

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=1]{kissings}}
	\caption{Two representations of a pair of kissing walks.}
	\label{fig:kissings}
\end{figure}
The left hand side of the picture is as in \fref{fig:kissing}, and the right hand side is simply a different representation of it.
On the surface~$\surface_{\bar Q}$, the dual dissections~$\dissection$ and~$\dissection\dual$ are as on \fref{fig:kissingVSCrossing} (on the right hand side).

@@

\begin{remark}
Note that our construction of the surface~$\surface_{\bar Q}$ implies that the vertices of~$\bar Q$ are naturally identified with the intersection points of the dual edges~$a \in \dissection_{\bar Q}$ and~$a\dual \in {\dissection_{\bar Q}}\dual$.
If we force all $\dissection$-accordions (or $\dissection\dual$-slaloms) to cross the edges of~$\dissection$ and~$\dissection\dual$ only at these intersection points, then an $\dissection$-accordion (or $\dissection\dual$-slalom) is really seen a walk on~$\bar Q$.
\end{remark}

We conclude with an important example of facets of the non-kissing and non-crossing complexes.

\begin{example}
\label{exm:peakDeepFacets}

\vincent{TODO: Describe here the peak and deep facets in terms of the surface.}
\end{example}

@@

 \para{Geometric proof}
 Let~$F$ be a face of the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$.
 Its elements are pairwise compatible~$\dissection$-accordions on the surface~$\surface_{\bar Q}$.
 Thus~$F$ can be viewed as a partial tagged triangulation of~$\surface_{\bar Q}$, by letting the~$\dissection$-accordions circling a puncture in (say) clockwise order be seen as ``notched'' arcs to said puncture, and the ones circling it in counter-clockwise orientation be ``un-notched'' arcs.
 It is known (see, for instance, \cite{FominShapiroThurston}) that any such partial tagged triangulation can be embedded into a full triangulation, and that triangulations contain a finite number of arcs.
 Thus~$F$ is finite, and is contained in a finite maximal face.
\end{proof}

@@

\para{Combinatorial proof}
Consider a maximal non-kissing face~$F \in \NKC$ and an arrow~$\alpha$ in~$c$.
Let~$\beta, \gamma \in Q_1\blossom$ be such that~$t(\alpha) = s(\beta) = s(\gamma)$ and~$\alpha\beta \notin I$ while~$\alpha\gamma \in I$ (in other words, $\beta$ is the next arrow along~$c$ after~$\alpha$ and~$\gamma$ is the other outgoing arrow at~$t(\alpha)$).
Consider the walk~$\mu \eqdef \distinguishedWalk{\gamma}{F}$ and decompose it as~$\mu = \mu_1 \gamma \mu_2$ according to this occurence of~$\gamma$ on~$\mu$.
Consider the walk~$\nu \eqdef \mu_2^{-1} \gamma^{-1} c^\infty$.
If~$\nu$ is contained in~$F$, then we have found a walk of the desired form (remember that walks are considered as undirected).
Otherwise, there exists a walk~$\lambda$ of~$F$ preventing~$\nu$ to belong to~$F$.
Let~$\sigma$ be a common substring of~$\nu$ and~$\lambda$ where they kiss.
Since~$\lambda$ must be compatible with~$\mu$ and with the straight walk~$c^\infty$, the substring~$\sigma$ must contain~$t(\alpha) = s(\beta)$.
If~$\lambda$ contains neither~$\alpha$ nor~$\beta$, then it contains~$\gamma$ and enters $\nu$ at~$s(\gamma)$.
Since $\lambda$ kisses~$\nu$, it also enters~$\mu_2$, contradicting the maximality of~$\mu$ for~$\prec_\gamma$.
Therefore, $\lambda$ contains either~$\alpha$ or~$\beta$.
We have thus found a bending walk~$\lambda \in F$ containing an arrow~$\alpha$ or~$\beta$ of~$c$, say~$\alpha$.
We can now moreover assume that~$\lambda = \distinguishedWalk{\alpha}{F}$.
If~$\lambda$ starts by~$^\infty c$, or ends by~$c^\infty$, then we have found a walk of the desired form.
Otherwise, decompose this walk as~$\lambda = \lambda_1 \alpha \lambda_2$ and consider the walk~$\omega = \lambda_1 c^\infty$.
We claim that~$\omega$ belongs to~$F$.
Otherwise, there exists a walk~$\rho$ preventing~$\omega$ to belong to~$F$.
Let~$\tau$ be a common substring of~$\omega$ and~$\rho$ where they kiss.
Since~$\omega$ must be compatible with~$\lambda$ and with the straight walk~$c^\infty$, the substring~$\tau$ must strictly contain the substring of~$\lambda$ along the cycle~$c$.
It follows that $\rho$ leaves~$\omega$ when it is on the cycle~$c$, thus with an incoming arrow, contradicting the maximality of~$\lambda$ for~$\prec_\alpha$.

\para{Geometric proof}
The statement of the lemma translates on~$\surface_{\bar Q}$ as follows: for any facet~$F$ of the non-crossing complex~$\NCC$ and any puncture of the dissection~$\dissection_{\bar Q}$, there exists a curve in~$F$ which goes to the puncture.
Assume that there is a puncture with no curves of~$F$ going to it.
Construct a curve~$\gamma$ as follows:~$\gamma$ first unwinds from the puncture, then~$\gamma$ leaves the local picture around the puncture by turning around another marked point (marked~$X$ on the picture below).

@@

%\begin{proposition}
%\label{prop:gcdVectorsSurface}
%For any walk~$\omega$ in any non-kissing facet~$F \in \NKC$,
%\begin{itemize}
%\item the~$\b{d}$-vectors of the walk~$\omega$, of the $\dissection$-accordion~$\curveof(\omega)$, and of the $\dissection\dual$-slalom~$\curveof(\omega)$ coincide,
%\item the~$\b{g}$-vectors of the walk~$\omega$, of the $\dissection$-accordion~$\curveof(\omega)$, and of the $\dissection\dual$-slalom~$\curveof(\omega)$ coincide,
%\item the~$\b{c}$-vectors of the walk~$\omega$ in~$F$, of the $\dissection$-accordion~$\curveof(\omega)$ in~$\curveof(F)$ and of the $\dissection\dual$-slalom~$\curveof(\omega)$ in~$\curveof(F)$ concide.
%\end{itemize}
%\end{proposition}
%
%\begin{proof}
%Definitions~\ref{def:cVectorsAccordions} and~\ref{def:gVectorsSlaloms} are immediate translations in terms of curves on the surface~$\surface_{\bar Q}$ of Definitions~\ref{def:cVectorsWalks} and~\ref{def:gVectorsWalks} respectively.
%\end{proof}
